High performance computing for global optimization problems

### 2018-05-18 2018/ITNT/ITNT2018.tex
Evolvent figures
Tables and figures numbers are corrected

The use of Peano curve $y(x)$ 
\[
\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq 
N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
\]
unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.  To implement this method of dimensionality reduction a 
numerically constructed curve (\textit{evolvent}) is used. The evolvent is 
$2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ 
metric, where $m$ is an evolvent construction parameter. Problems of 
numerical construction of the evolvents and the corresponding theory are 
considered in detail in \cite{Strongin2000}.

@@

The level lines of the two-dimensional subproblems based on the functions 
$f(y_1,y_2)$ and $g(y_3,y_4)$ are shown in figure~1. The feasible domains are highlighted by color. One can see that 
the subproblem (a) has more complex structure as compared to the subproblem 
(b). At the same time, computing the values of the function $f(y_1,y_2)$ is 
more time-consuming than of $g(y_3,y_4)$.

\begin{figure}[ht]
\begin{minipage}[h]{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}[h]{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{GKLS_04.png} \\ (b)}
\end{minipage}
\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple 
subproblems}
\end{figure}

@@

Let us conduct the first experiment by solving the \textit{Simple} and \textit{Hard} problem 
series on a single node employing both available CPUs in full (i. e. $p=16$ 
cores available). In table 1 the averaged time (in seconds) 
required to solve the problems of the series is presented.

\begin{table}[ht]
\center{{\bf Table 1.} Average time for solving the problem on one node}
\begin{center}
\begin{tabular}{cc}
\hline
Problems & Time (sec) \\
\hline
\textit{Hard} & 52  \\
\textit{Simple} & 51 \\
\hline
\end{tabular}
\end{center}
\end{table}

Let us conduct the next experiment employing $p$ nodes each including three 
graphic accelerators. Thus, at $p=32$ total 96 GPU accelerators were employed 
each including 2688 CUDA cores. In tables 2 and 3 the average time and the speedup with respect to the run 
on a single node are presented.

\begin{table}[ht]
\center{{\bf Table 2.} Average time for solving the problem on $p$ nodes}
\begin{center}
\begin{tabular}{cccc}
\hline
Problems & $p = 8$ & $p=16$ & $p=32$ \\
\hline
\textit{Hard} & 11.51 & 5.53 & 0.54 \\
\textit{Simple} & 2.04 & 1.50 & 0.47\\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[ht]
\center{{\bf Table 3.} Speedup with respect to one node}
\begin{center}
\begin{tabular}{cccc}
\hline
Problems & $p = 8$ & $p=16$ & $p=32$ \\
\hline
\textit{Hard} & 5 & 9 & 96\\
\textit{Simple} & 25 & 34& 109\\
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

The use of Peano curve $y(x)$ 
\[
\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq 
N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
\]
unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.  To implement this method of dimensionality reduction a 
numerically constructed curve (\textit{evolvent}) is used. The evolvent is 
$2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ 
metric, where $m$ is an evolvent construction parameter. Problems of 
numerical construction of the evolvents and the corresponding theory are 
considered in detail in \cite{Strongin2000}. Examples of the evolvent with different $m$ in two dimensions are given in figure~\ref{fig:0}.

\begin{figure}[t]
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
\end{minipage}
%\hfill
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (c)}
\end{minipage}
\caption{Evolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
\label{fig:0}
\end{figure}

@@

The level lines of the two-dimensional subproblems based on the functions 
$f(y_1,y_2)$ and $g(y_3,y_4)$ are shown in figure~\ref{fig:2}. The feasible domains are highlighted by color. One can see that 
the subproblem (a) has more complex structure as compared to the subproblem 
(b). At the same time, computing the values of the function $f(y_1,y_2)$ is 
more time-consuming than of $g(y_3,y_4)$.

\begin{figure}[ht]
\begin{minipage}[h]{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}[h]{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{GKLS_04.png} \\ (b)}
\end{minipage}
\caption{The level lines of (a) hard and (b) simple subproblems}
\label{fig:2}
\end{figure}

@@

Let us conduct the first experiment by solving the \textit{Simple} and \textit{Hard} problem 
series on a single node employing both available CPUs in full (i.e. $p=16$ 
cores available). In table \ref{tab:1} the averaged time (in seconds) 
required to solve the problems of the series is presented.

\begin{table}
\caption{Average time for solving the problem on one node}
\label{tab:1}
\begin{center}
\begin{tabular}{cc}
\hline
Problems & Time (sec) \\
\hline
\textit{Hard} & 52  \\
\textit{Simple} & 51 \\
\hline
\end{tabular}
\end{center}
\end{table}

Let us conduct the next experiment employing $p$ nodes each including three 
graphic accelerators. Thus, at $p=32$ total 96 GPU accelerators were employed 
each including 2688 CUDA cores. In tables \ref{tab:2} and \ref{tab:3} the average time and the speedup with respect to the run 
on a single node are presented.

\begin{table}
\caption{Average time for solving the problem on $p$ nodes}
\label{tab:2}
\begin{center}
\begin{tabular}{cccc}
\hline
Problems & $p = 8$ & $p=16$ & $p=32$ \\
\hline
\textit{Hard} & 11.51 & 5.53 & 0.54 \\
\textit{Simple} & 2.04 & 1.50 & 0.47\\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}
\caption{Speedup with respect to one node}
\label{tab:3}
\begin{center}
\begin{tabular}{cccc}
\hline
Problems & $p = 8$ & $p=16$ & $p=32$ \\
\hline
\textit{Hard} & 5 & 9 & 96\\
\textit{Simple} & 25 & 34& 109\\
\hline
\end{tabular}
\end{center}
\end{table}
